---
output: 
  html_document: 
    keep_md: yes
---
# Reproducible Research: Project 1
# Load and process the data

# 1. Load the data 
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```
```{r, results='markup', warning=TRUE, message=TRUE}

activityData <- read.csv('activity.csv')
```
# 2. Process/transform interval data

```{r}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```
-----
# A. What is mean total number of steps taken per day?

# 1. total number of steps taken per day.

```{r}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```

# 2. Histogram of the total number of steps taken each day

```{r}
hist(stepsByDay,col="red",breaks=20,main="Total steps per day",xlab="Steps per day")
```

# 3. Mean and median total number of steps taken per day

```{r}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```
* Mean: `r stepsByDayMean` 
* Median: `r stepsByDayMedian`

-----

# B. What is the average daily activity pattern?

```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

# 1. Time series plot

```{r}

plot(x = averageStepsPerTimeBlock$interval ,y = averageStepsPerTimeBlock$meanSteps,type = "l", col = "blue", xlab="Intervals", ylab = "Average Daily Steps")

```

## 2. The 5-minute interval on average across all the days in the dataset that contains the maximum number of steps?

```{r}
MaxSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMaxSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[MaxSteps,'interval'])
```

* Maximum Steps at: `r timeMaxSteps`




# C. Imputing Missing Values:

# 1. The total number of missing values in the dataset 

```{r}
numMissingValues <- length(which(is.na(activityData$steps)))
```

* Number of missing values: `r numMissingValues`

# 2. Devise a strategy for filling in all of the missing values in the dataset.
# 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```

# 4. Histogram of the total number of steps taken each day & Calculate mean and median 

```{r}
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)

hist(stepsByDayImputed,col="red",breaks=20,xlab="Total steps per day (Imputed)")

```

```{r}
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
* Mean (Imputed): `r stepsByDayMeanImputed`
* Median (Imputed):  `r stepsByDayMedianImputed`

----

# D. Are there differences in activity patterns between weekdays and weekends?

# 1. Create a new factor variable in the dataset with two levels; "weekday' and "weekend"", indicating whether a given date is a weekday or weekend day.
```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

# 2. Panel plot containing a time series plot
```{r}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5 Minute interval") + 
    ylab("Avarage number of steps")
```